### 2020-06-05 writeup/intro.tex
intro quick pass

Second, we explain the performance of multi-task learning by varying the three components above in isotropic and covariate shifted settings.
We describe three simulations in Figure \ref{fig_model_shift_phasetrans} for two linear regression tasks and measure the loss of the single-task estimator minus the multi-task estimator.
%We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
%Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
\squishlist
	\item We provide conditions to predict the effect of transfer in the isotropic setting (Proposition \ref{prop_dist_transition}).
	Our result predicts most of the empirical observations in Figure \ref{fig_model_shift} correctly.
%	It is crucial that the concentration result in Theorem \ref{lem_cov_shift_informal} is sufficiently precise so that we can explain the transition phenomena in Figure \ref{fig_model_shift} and \ref{fig_size}.
	The unexplained observations are caused by an error term from the bias.
	We discuss these in Section \ref{sec_insight}.
	\item We provide a similar condition that explains a curious phenomenon in Figure \ref{fig_size}, namely adding more labeled data from the source task does not always improve performance (Proposition \ref{prop_data_size}).
	%Theorem \ref{lem_cov_shift_informal} allows us to compare MTL performance under different covariate shifts.
	\item We study a special case of having the same task models and show that MTL  performs best without covariate shift (Proposition \ref{prop_covariate}).
	For this special case, the bound of Lemma \ref{lem_cov_shift_informal} matches the empirical curve in Figure \ref{fig_covariate} accurately.
\squishend
%In Section \ref{sec_insight}, we consider three components including task similarity, data size and covariate shift for a simplified isotropic setting of two tasks.
%We measure task similarity by how small is the distance between $\beta_1$ and $\beta_2$.
%Using our tool, we explain a transition from positive to negative transfer as task similarity decreases.
%		Furthermore, we show that negative transfer is more likely to occur when the source task labels are particularly noisy.
%		In Section \ref{sec_validate}, we validate the observation on text and image classification tasks.
%	In , we provide the trade-off between $\norm{\beta_1 - \beta_2}^2$ and a certain function $\Phi(\rho_1, \rho_2)$ to determine the type of transfer.
%We show that increasing the data size of the source task does not always improve performance for the target task in multi-task learning.
Along the way, we explain a key empirical finding of Taskonomy by Zamir et al. \cite{ZSSGM18}, by analyzing the benefit of MTL for reducing labeled data to achieve comparable performance to STL.
%We show that covariate shift, measured by $\Sigma_1^{1/2}\Sigma_2^{-1/2}$, is another cause for suboptimal performance for $\hat{\beta}_t^{\MTL}$.
%		We show that as $n_1 / n_2$ becomes large, having no covariate shift between the source and target tasks yields the optimal performance for the target task.
%		On the other hand, when $n_1 / n_2$ is small, there are counter examples where having the same covariance matrix is not necessarily the optimal choice.

Finally, we connect our theory to practical problems of interest.
We provide a single-task based metric to predict positive or negative transfer in multi-task learning.
While it is not well understood when multi-task learning provides positive transfer, we show that the STL results can help indicate and understand MTL results on ChestX-ray14 \cite{chexnet17} and sentiment analysis datasets \cite{LZWDA18}.
Inspired by our observation that adding more source task data does not always improve performance, we propose to train multi-task models by incrementally adding source task data for predicting a target task.
We show that our training schedule reduces the computational cost by $55\%$ compared to baseline multi-task training on the sentiment analysis dataset while keeping the accuracy the same.
Last, we validate our three theoretical findings on sentiment analysis tasks.

>>>>

Second, we explain the performance of multi-task learning by varying the three components above in isotropic and covariate shifted settings.
We describe three simulations in Figure \ref{fig_model_shift_phasetrans} for two linear regression tasks and measure the loss of the single-task estimator minus the multi-task estimator.
%We observe that negative transfer occurs as (a) \textit{task similarity}: tasks become more different; (b) \textit{data size}: source/target data size increases.
%\textbf{Task similarity:}
%\textbf{Data sizes:}
%\textbf{Covariate shift:}
%Furthermore, MTL performance is negatively affected when (c) \textit{covariate shift}: the covariance matrices of the two tasks become more different.
\squishlist
	\item We provide conditions to predict the effect of transfer in the isotropic setting (Proposition \ref{prop_dist_transition}).
	Our result predicts most of the empirical observations in Figure \ref{fig_model_shift} correctly.
%	It is crucial that the concentration result in Theorem \ref{lem_cov_shift_informal} is sufficiently precise so that we can explain the transition phenomena in Figure \ref{fig_model_shift} and \ref{fig_size}.
	The unexplained observations are caused by an error term from the bias.
	We discuss these in Section \ref{sec_insight}.
	\item We provide a similar condition that explains a curious phenomenon in Figure \ref{fig_size}, namely adding more labeled data from the source task does not always improve performance (Proposition \ref{prop_data_size}).
	%Theorem \ref{lem_cov_shift_informal} allows us to compare MTL performance under different covariate shifts.
	\item We study a special case of having the same models and show that MTL  performs best without covariate shift (Proposition \ref{prop_covariate}).
	Our theoretical bound matches the empirical curve in Figure \ref{fig_covariate}.
\squishend
%In Section \ref{sec_insight}, we consider three components including task similarity, data size and covariate shift for a simplified isotropic setting of two tasks.
%We measure task similarity by how small is the distance between $\beta_1$ and $\beta_2$.
%Using our tool, we explain a transition from positive to negative transfer as task similarity decreases.
%		Furthermore, we show that negative transfer is more likely to occur when the source task labels are particularly noisy.
%		In Section \ref{sec_validate}, we validate the observation on text and image classification tasks.
%	In , we provide the trade-off between $\norm{\beta_1 - \beta_2}^2$ and a certain function $\Phi(\rho_1, \rho_2)$ to determine the type of transfer.
%We show that increasing the data size of the source task does not always improve performance for the target task in multi-task learning.
Along the way, we explain a key empirical finding of Taskonomy by Zamir et al. \cite{ZSSGM18}, by analyzing the benefit of MTL for reducing labeled data to achieve comparable performance to STL.
%We show that covariate shift, measured by $\Sigma_1^{1/2}\Sigma_2^{-1/2}$, is another cause for suboptimal performance for $\hat{\beta}_t^{\MTL}$.
%		We show that as $n_1 / n_2$ becomes large, having no covariate shift between the source and target tasks yields the optimal performance for the target task.
%		On the other hand, when $n_1 / n_2$ is small, there are counter examples where having the same covariance matrix is not necessarily the optimal choice.

Our study also leads to several algorithmic consequences with practical interest.
First, we show that single-task learning results can help predict positive or negative transfer for multi-task learning.
We validate this observation on ChestX-ray14 \cite{chexnet17} and sentiment analysis datasets \cite{LZWDA18}.
Second, we propose to train multi-task models by incrementally adding source task data for predicting a target task.
This is inspired by our observation in Figure \ref{fig_size} where adding more source task data helps initially, but hurts eventually.
Using our incremental training schdule, we reduce the computational cost by $55\%$ compared to baseline multi-task training over six sentiment analysis dataset while keeping the accuracy the same.
Third, we provide a fine-grained insight on a covariange alignment procedure proposed in \cite{WZR20}.
We show that the alignment procedure provides more significant improvement when the source/target sample ratio is large.
Finally, we validate our three theoretical findings on sentiment analysis tasks.
